Bayesian phylogenetic estimation of fossil ages reveals remarkable consistency of rocks and clocks

### 2015-08-26 fossilDating.tex
adding canid analysis

\Mrelaxed{} was the best-fitting model for this data set according to the analysis of \textcite{gavryushkina2015bayesian}. As with \Mstrict{} this model produced phylogenetic age estimates that were very concordant with the geological age ranges of the fossils (Figure \ref{fig:phyloAgeVsGeoAge}b). The median error was 2.1 Myr across all 36 fossils. Again, as with \Mstrict{}, none of the fossils exhibited strong evidence (i.e. $|\text{log BF}| > 3.0$) that the phylogenetic age was inconsistent with the geological age range assuming a uniform prior from $(0,160)$ Mya. However if we consider the posterior probability that the fossil is in the geological age range then five of the 36 fossils had a posterior probability $< 0.05$ for \Mrelaxed{}, suggesting low posterior support for the phylogenetic age being within the palaeontological age range. These five fossils were {\em Madrynornis mirandus}, {\em Paraptenodytes antarcticus}, {\em Perudyptes devriesi}, {\em Sphenicus muizoni} and {\em Waimanu manneringi} with posterior probabilities that the phylogenetic age is in the palaeontological range of 0.024, 0.016, 0.048, 0.007, 0.037 respectively. All other fossils has posterior probabilities of $> 0.05$ of their age being in the palaeontological range. Again the absolute discrepancy in the ages are quite moderate for the five fossils with low posterior probabilities, with {\em M. mirandus}: 6.5Myr vs 10Myr (phylogenetic age versus palaeontological age), {\em P. antarcticus}: 28.1 vs 22, {\em P. devriesi}: 49.0 vs 40, {\em S. muizoni}: 5.1 vs 9.1 and {\em W. manneringi}: 56.8 vs 61.05. A summary of all the individual estimates are tabulated in Table \ref{fossilTable8}. The individual marginal posterior distributions of phylogenetic age under \Mrelaxed{} and corresponding geological range are shown in Figures \ref{hist8_younger} and \ref{hist8_older}.

@@

>>>>

\Mrelaxed{} was the best-fitting model for this data set according to the analysis of \textcite{gavryushkina2015bayesian}. As with \Mstrict{} this model produced phylogenetic age estimates that were very concordant with the geological age ranges of the fossils (Figure \ref{fig:phyloAgeVsGeoAge}b). The median error was 2.1 Myr across all 36 fossils. Again, as with \Mstrict{}, none of the fossils exhibited strong evidence (i.e. $|\text{log BF}| > 3.0$) that the phylogenetic age was inconsistent with the geological age range assuming a uniform prior from $(0,160)$ Mya. However if we consider the posterior probability that the fossil is in the geological age range then five of the 36 fossils had a posterior probability $< 0.05$ for \Mrelaxed{}, suggesting low posterior support for the phylogenetic age being within the palaeontological age range. These five fossils were {\em Madrynornis mirandus}, {\em Paraptenodytes antarcticus}, {\em Perudyptes devriesi}, {\em Sphenicus muizoni} and {\em Waimanu manneringi} with posterior probabilities that the phylogenetic age is in the palaeontological range of 0.024, 0.016, 0.048, 0.007, 0.037 respectively. All other fossils has posterior probabilities of $> 0.05$ of their age being in the palaeontological range. Again the absolute discrepancy in the ages are quite moderate for the five fossils with low posterior probabilities, with {\em M. mirandus}: 6.5Myr vs 10Myr (phylogenetic age versus palaeontological age), {\em P. antarcticus}: 28.1 vs 22, {\em P. devriesi}: 49.0 vs 40, {\em S. muizoni}: 5.1 vs 9.1 and {\em W. manneringi}: 56.8 vs 61.05. A summary of all the individual estimates are tabulated in Table \ref{fossilTable8}. The individual marginal posterior distributions of phylogenetic age under \Mrelaxed{} and the corresponding geological range are shown in Figures \ref{hist8_younger} and \ref{hist8_older}.

@@

\begin{figure}
\includegraphics{canids/1c_phyloAgeVsGeoAge.pdf}
\caption{\label{fig:canidMorphVsGeo}
The Bayesian phylogenetic age estimate (median and credible interval of marginal posterior) for each of the 123 canid fossils plotted against their palaeontological age point estimates, under a strict morphological clock model \Mstrict{}. The Bayesian estimates are represented by the median of the marginal posterior distribution and the upper and lower limits of the 99\% HPD interval. Blue line shows the $x=y$. If the vertical line doesn't cross $x=y$, then the geological age is not in the credible interval of the phylogenetic age. The 16 fossils for which this is the case are labelled.}
\end{figure}
